---
title: "CKME 136 - Mount Rainier Climbing Dataset"
author: "Nicole Smith"
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
---




```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```





```{r}
#first, load the dataset
rainier<-read.csv("C:/Users/nicol/OneDrive/Desktop/rainier.csv")
summary(rainier)

```


```{r}
#remove rows with more than 4 nas.
nrow(rainier)

keep<-rowSums(is.na(rainier)) <4
rainierclean<-rainier[keep, ]

nrow(rainierclean)

#count how many nas remain
sum(is.na(rainierclean))

table(is.na(rainierclean))

sum(is.na(rainierclean$Solar_Radiation))


#now remove all remaining nas
rainierclean<- rainierclean[complete.cases(rainierclean), ]

nrow(rainierclean)

#now we need to remove any rows where the success percentage is greater than 1, since that is not logical.

rainier.clean<-rainierclean[(rainierclean$Success_Percentage<=1), ]

nrow(rainier.clean)

summary(rainier.clean)

#now, we have a clean dataset.
```



```{r}
#correlation


cor(rainier.clean$Success_Percentage, rainier.clean$Battery_V)

cor(rainier.clean$Success_Percentage, rainier.clean$Temp)

cor(rainier.clean$Success_Percentage, rainier.clean$Humidity)

cor(rainier.clean$Success_Percentage, rainier.clean$Wind_Speed)

cor(rainier.clean$Success_Percentage, rainier.clean$Wind_Direction)

#wind speed and temp have the highest effect on success rates, and battery voltage has the lest. for the linear model i will use temp, humidity, wind speed and wind direction.
```


```{r}
#regression model
regmodel<-lm(rainier.clean$Success_Percentage ~ rainier.clean$Temp + rainier.clean$Humidity + rainier.clean$Wind_Speed + rainier.clean$Wind_Direction)

regmodel

#success rate=0.2921594+0.0007881(humidity)-0.0002947(wind direction)+0.0053524(temperature)-0.0044135(wind speed)
```


```{r}
#look at which route has the best probability of success
agg1<-aggregate(Success_Percentage ~ Route, rainier.clean, mean)
agg1

agg2<-aggregate(Success_Percentage ~ Route, rainier.clean, sum)
agg2

#although the tahoma cleaver has the highest average success percentage, of 100%, there is only 1 attempt, so we do not have enough evidence to conclude that this route has the highest success percentage.
```



```{r}
#export to excel for sql analysis
#write.xlsx(rainier.clean, "C:/Users/nicol/OneDrive/Desktop/rainierclean.xlsx")
```

